---
title: "AGRON 920 - Final Project"
output: html_document
date: "2025-05-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#install.packages("soilDB")
#remotes::install_github("mikejohnson51/climateR")
#remotes::install_github("rspatial/geodata")
#install.packages("tidygeocoder")
#install.packages("vip")
#install.packages("aqp")
#install.packages("rnaturalearth")
```

## Libraries

```{r}
library(soilDB)
library(tidyverse)
library(aqp)
library(rnaturalearth)
library(climateR)
library(readxl)
library(raster)
library(sf)
library(tidygeocoder)
library(NbClust)
library(factoextra)
library(spdep)
library(ade4)
library(adespatial)
library(geocmeans)
library(terra)
library(tidymodels)     
library(doParallel)     
library(tidyverse)      
library(ranger)         
library(finetune)     
library(readxl)         
library(patchwork)      
tidymodels_prefer()
library(vip)
library(lwgeom)
library(mapview)
```

```{r}
palette<-c("#3b0001","#be0003","#ff4903","#ffdc00","#559d03" ,"#075e07",'#161f0c')
rdylgn<-colorRampPalette(palette)
```


## Study area 

```{r fig.width=4, fig.height=5}

world <- rnaturalearth::ne_countries(scale = 'small', returnclass = 'sf')

arg_location <- ggplot() + 
  geom_sf(data = world) + 
  geom_sf(data = argentina_sf,color = 'grey40', fill = 'grey40') +
  coord_sf(xlim = c(-165,-30),crs = 'epsg:4326') +
   theme() + 
   labs(subtitle = 'Study area location')

ggsave(plot = arg_location,filename = '../plots/arg_location.png',width = 4,height = 5)

```


```{r}
argentina_sf <- rnaturalearth::ne_states() %>%
  filter(admin == 'Argentina')

argentina_map <- ggplot() + 
  geom_sf(data = argentina_sf) +
  geom_sf(data = cordoba_sf,fill = '#343a40') +
  labs(x = 'Longitude', y = 'Latitude')

provinces <- argentina_sf %>%
  filter(name %in% c('Entre Ríos','Salta','Chaco','Corrientes','Buenos Aires',
  'Tucumán','Córdoba','La Pampa','Santa Fe','Santiago del Estero','San Luis'))


```

## 2. Data collection

### 2.1 Crop data

```{r}
crop_data <- read_xlsx('../data/data_ready_23_24.xlsx')

crop_data <- crop_data %>%
  mutate(
    address = paste(Trial_Site, Province, "Argentina", sep = ", ")
  ) %>%
  group_by(address) %>%
  mutate(address_id = cur_group_id(),
         Env_index = mean(Rendimiento)) %>%
  ungroup()


crop_data <- geocode(.tbl =  crop_data, 
                   address = address, method = "osm", verbose = TRUE) %>%
  filter(is.na(long) == FALSE) %>%
  st_as_sf(coords = c('long','lat')) %>%
  st_set_crs(value = 'EPSG:4326:') 

crop_data <- crop_data %>%
  dplyr::select(c(Province,Trial_Site,GM,Cultivar, Empresa,Rendimiento,Env_index, geometry ))

```

```{r}
study_area_plot <- crop_data %>%
  ggplot() + 
  geom_sf(data = argentina_sf) +
  geom_sf(aes(color = Env_index),size = 2) +
  geom_sf_text(aes(label = Trial_Site),check_overlap = TRUE,nudge_y = 0.25) +
  
  geom_sf(data = provinces,fill = NA,linewidth = 1.2) +
  scale_color_gradientn(colours = rdylgn(100))  +
  coord_sf(ylim = c(-42,-22),xlim = c(-70,-55)) + 
  ylab("Latitude") + xlab("Longitude") + 
  labs(title = 'Experiment distribution across study area',
       color = 'Environmental \nIndex')


ggsave(plot = study_area_plot,filename = '../plots/study_area.png',width = 8,height = 6)

study_area_plot
```


### 2.2 Climate data 

```{r}
weather_vars <- climateR::getTerraClimNormals(AOI = argentina_sf,
                              varname = c('aet',"ppt",'tmax','tmin',"vpd"),
                              month = c(1,2,3,4,11,12))

plot(weather_vars$tmax)
```

### 2.3 Soil data

```{r}
# soil_data <- soilDB::fetchSoilGrids(x = argentina_sf,depth_intervals = "0-5",
#                        variables = c("soc", "bdod", "ocd", "ocs"), 
#                        filename = '../data/soil_data.tif',
#                        grid = TRUE)


soil_data <- geodata::soil_world(stat="mean",
                                 var=c("soc","bdod",
                                       "sand",'silt','phh2o'),
                           depth=5, path=tempdir())


soil_data_crop <- crop(soil_data, argentina_sf)
soil_data_mask <- mask(soil_data_crop, argentina_sf)

plot(soil_data_mask)
```


## 3.Data integration


### 3.1 Grid creation

```{r}
argentina_grid <- provinces %>%
  st_set_crs(st_crs("epsg:4326")) %>%
  st_transform(32720) %>%
  st_make_grid(cellsize = 10000, what = "centers") %>% #cellsize = 10 meters
  st_as_sf() %>% 
  st_intersection(st_transform(st_as_sf(provinces),crs = 32720)) %>%
  st_set_crs(st_crs("epsg:32720")) %>%
  mutate(X_proj = st_coordinates(.)[,1],
         Y_proj = st_coordinates(.)[,2]) %>%
  st_transform(4326) %>%
  mutate(X = st_coordinates(.)[,1],
         Y = st_coordinates(.)[,2]) %>%
  rowid_to_column() %>%
  dplyr::select(c(rowid,name,X,Y,X_proj,Y_proj))

plot(argentina_grid)
```


```{r fig.height=8, fig.width=5}
ggplot() + 
  geom_sf(data = argentina_grid,size=0.01, alpha=0.4) + 
  geom_sf(data = argentina_sf, fill=NA, color='black',linewidth = 1, alpha=0.1) + 
  coord_sf(crs = 3857) +
  labs(x = 'Longitude', y = 'Latitude', 'Grid point across Kansas')
```


```{r message=FALSE}
df_grid <- data.frame(argentina_grid, 
           raster::extract(weather_vars$aet, argentina_grid),
           raster::extract(weather_vars$tmax, argentina_grid),
           raster::extract(weather_vars$ppt, argentina_grid),
           raster::extract(weather_vars$tmin, argentina_grid),
           raster::extract(weather_vars$vpd, argentina_grid),
           
           raster::extract(soil_data_mask$`soc_0-5cm`, argentina_grid),
           raster::extract(soil_data_mask$`bdod_0-5cm`, argentina_grid),
           raster::extract(soil_data_mask$`sand_0-5cm`, argentina_grid),
           raster::extract(soil_data_mask$`silt_0-5cm`, argentina_grid),
           raster::extract(soil_data_mask$`phh2o_0-5cm`, argentina_grid)) %>%
  dplyr::select(-c(ID.1,ID.2,ID.3,ID.4,ID.5,ID.6,ID.7,ID.8,ID.9)) %>%
  na.exclude() %>%
  rename_with(.fn = ~ map_chr(., function(col_name) {
    str_replace(
      col_name,
      "^([a-zA-Z0-9]+)_\\d{4}\\.(\\d{2})\\.\\d{2}_\\d{8}$",
      "\\1_\\2"
    )
  }))



cols_filtered <- names(df_grid) %>%
  keep(~ str_detect(., paste0("_(", paste(sprintf("%02d", c(1, 2, 3, 4, 11, 12)) , collapse = "|"), ")$")))

df_grid.X <- df_grid %>%
  dplyr::select(X_proj,Y_proj, all_of(cols_filtered),soc_0.5cm,bdod_0.5cm,sand_0.5cm,silt_0.5cm,phh2o_0.5cm)

```




## 4. Feature engineering

### 4.1 Principal component analysis

```{r}
Biplot_sPCA <- function(sPca, title){

PCA <- cbind(sPca$c1[,1], sPca$c1[,2])
PCA <- tibble::rownames_to_column(as.data.frame(PCA),
                                  var = "Variables")
colnames(PCA) <- c("variables","PC1","PC2")
rownames(PCA) <- rownames(sPca$c1)

angle <- seq(-pi, pi, length = 50) 
df <- data.frame(x = sin(angle), y = cos(angle)) 

angle <- seq(-pi, pi, length = 50) 
df <- data.frame(x = sin(angle), y = cos(angle)) 

pc1_var_exp <- as.double(c(as.vector(sPca$eig/sum(sPca$eig)))[1])*100
pc2_var_exp <- as.double(c(as.vector(sPca$eig/sum(sPca$eig)))[2])*100

pplot_pca1 <- ggplot() +
  geom_path(aes(x, y), data = df, colour="grey70") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  geom_hline(yintercept = 0, linetype = "dashed") +
  xlab(paste0("PC1 (", round(pc1_var_exp,2),"%)")) + 
  ylab(paste0("PC2 (", round(pc2_var_exp,2),"%)")) +
  geom_segment(data=PCA, show.legend = TRUE,size= 1.2, lineend = "round", linejoin = "round",  
               aes(x = 0, y = 0, xend = PC1, yend = PC2, colour = as.numeric(PC1)), alpha=0.4, 
               arrow = arrow(length = unit(2, 'mm'), type = "open"),
               color = "orange") +

  ggrepel::geom_label_repel(data=PCA, aes(x=PC1,y=PC2,
                           label=rownames(PCA)), colour = 'black', fill = 'white',
                           size=3, direction="both", max.overlaps = 350) +
  labs(color=PCA$PC1, title = "", subtitle = title)  +
   theme(aspect.ratio= 1) + 
   theme_bw(base_size = 15) 
   
return(pplot_pca1)
}

```

```{r}
PCA_vars <- dudi.pca(df_grid.X[,-c(1:2)], 
                center=T, scale = TRUE, 
                scannf = FALSE, nf = 5)

coords  <- sp::coordinates(df_grid.X[,c(1:2)]) 
gri <- dnearneigh(coords,0,150000) 
lw2<- nb2listw(gri, style = "W")


sPCA_vars <- adespatial::multispati(PCA_vars, lw2, 
                                 scannf = F, nfposi = 5)

df_PCA <- sPCA_vars$li[,1:2]
colnames(df_PCA) <- c("CS1","CS2")
df_grid <-cbind(df_grid, df_PCA)

df_grid[,c('CS1','CS2')]
```

```{r fig.height=10, fig.width=10}
biplot_sPCA <- Biplot_sPCA(sPCA_vars, 'Biplot - sPCA')
biplot_sPCA
ggsave("../plots/biplot_sPCA.png",biplot_sPCA,
       width = 15,height = 15,units = "cm")
```


### 4.2 Spatial clustering analysis

```{r}
future::plan(future::multicore(workers = availableCores(constraints = "multicore")))
FCM2<- select_parameters.mc("SFCM", df_grid[,c('CS1','CS2')], 
                            k = 2, m = 1.3, alpha = 1.5, 
                            nblistw = lw2, lag_method='median')

FCM3 <- select_parameters.mc("SFCM", df_grid[,c('CS1','CS2')], 
                            k = 3, m = 1.3, alpha = 1.5, 
                            nblistw = lw2, lag_method='median')

FCM4 <- select_parameters.mc("SFCM", df_grid[,c('CS1','CS2')], 
                            k = 4, m = 1.3, alpha = 1.5, 
                            nblistw = lw2, lag_method='median')

FCM5 <- select_parameters.mc("SFCM", df_grid[,c('CS1','CS2')], 
                            k = 5, m = 1.3, alpha = 1.5, 
                            nblistw = lw2, lag_method='median')

FCM6 <- select_parameters.mc("SFCM", df_grid[,c('CS1','CS2')], 
                            k = 6, m = 1.3, alpha = 1.5, 
                            nblistw = lw2, lag_method='median')

FCM7 <- select_parameters.mc("SFCM", df_grid[,c('CS1','CS2')], 
                            k = 7, m = 1.3, alpha = 1.5, 
                            nblistw = lw2, lag_method='median')


FCM_indices <- rbind(as.numeric(FCM2),
                     as.numeric(FCM3),
                     as.numeric(FCM4),
                     as.numeric(FCM5),
                     as.numeric(FCM6),
                     as.numeric(FCM7))

FCM_indices <- t(FCM_indices)
FCM_indices <- FCM_indices[0:10,]

rownames(FCM_indices) <- c('Silhouette Index',
                           'Partition Entropy',
                           'Partition Coef.',
                           'Mod. Partit. Coef.',
                           'Xie-Beni index',
                           'Fukuyama-Sugeno index',
                           'Explained.inertia',
                           'spConsistency',
                           'spConsistency_05.5.',
                           'spConsistency_95.95.')

colnames(FCM_indices) <- c('C2','C3','C4','C5','C6','C7')

FCM_climate_cluster_indices <- tibble::rownames_to_column(as.data.frame(FCM_indices),
                                  var = "Indices")

FCM_climate_cluster_indices <- FCM_climate_cluster_indices %>%
                  tidyr::pivot_longer('C2':'C7', 
                    names_to = "Clusters", 
                    values_to = "value", names_repair = "minimal")


plot_index_cluster <- ggplot(FCM_climate_cluster_indices, aes(Clusters, value), group = Indices) + 
  geom_line(aes(colour = Indices,group = Indices)) +
  geom_point(aes(colour = Indices, group = Indices)) + 
  labs(title = "Optimal number of classes - Method Spatial Fuzzy c-means ",
       subtitle = "Clustering analysis using soil and climate data") + 
  facet_wrap(~Indices, scales = "free") 

ggsave("../plots/cluster_indexes.png",width = 12)  
plot_index_cluster

```

```{r}
future::plan(future::multicore(workers = availableCores(constraints = "multicore")))
FCM_cluster_grid <- SFCMeans(df_grid[,c('CS1','CS2')], 
                               lw2,
                               k = 6, 
                               m = 1.5, 
                               alpha = 0.3, 
                               standardize = FALSE,
                               lag_method = 'median',
                               verbose = FALSE)
```




```{r}

df_grid_clusters <- cbind(df_grid, as.integer(as.factor(FCM_cluster_grid$Groups))) %>%
  rename(clusters = `as.integer(as.factor(FCM_cluster_grid$Groups))`)


cluster_raster <- rast(df_grid_clusters[c('X_proj','Y_proj','clusters')], type = "xyz") %>%
  `crs<-`("EPSG:32720")  %>%
  project("EPSG:4326")


clusters_arg <- ggplot() +
  geom_sf(data = argentina_sf) + 
  geom_point(data = df_grid_clusters, aes(x = X,y = Y,color = as.factor(clusters)), size=0.5) +
  geom_sf(data = provinces,color = 'black', fill=NA, linewidth = 0.5,alpha = 0.5) +
  geom_sf(data = crop_data,color = 'black') + 
  scale_color_viridis_d(option = 'plasma') +
  coord_sf(ylim = c(-42,-22),xlim = c(-70,-55)) + 
  ylab("Latitude") + xlab("Longitude") + 
      labs(color= 'Clusters',title = "Agro-Environmental zones in Argentina",
       subtitle = "sPCA + Geo Fuzzy c-means")
  
ggsave(clusters_arg,filename = '../plots/clusters_arg.png',width = 6,height = 6)

clusters_arg
```

```{r}
df_summary_weather <- df_grid_clusters %>%
  dplyr::select(-c(name, X, Y, X_proj, Y_proj, ID, x,  
                   bdod_0.5cm, sand_0.5cm, soc_0.5cm, 
                   silt_0.5cm, phh2o_0.5cm, CS1, CS2)) %>%
  pivot_longer(
    cols = matches("^(aet|ppt|tmin|tmax|vpd)_"),
    names_to = c("variable", "mes"),
    names_sep = "_",
    values_to = "valor"
  ) %>%
  mutate(mes = as.integer(mes)) %>%
  filter(mes %in% c(1,2,3,4,11,12)) %>%
  group_by(rowid,clusters, variable) %>%
  reframe(
    resumen = case_when(
      variable == "ppt" ~ sum(valor, na.rm = TRUE),
      TRUE ~ mean(valor, na.rm = TRUE)
    ),
    .groups = "drop"
  ) %>%
  pivot_wider(
    names_from = variable,
    values_from = resumen
  ) %>%
  unnest() %>%
  pivot_longer(cols = -c(rowid, clusters, .groups),
               names_to = 'weather_vars',values_to = 'values') %>% 
  distinct()
```

```{r fig.height=5,fig.width=10}
plot_summary_weather <- df_summary_weather %>%
  dplyr::select(-c(rowid,`.groups`)) %>%
  mutate(clusters = as.factor(clusters)) %>%
  #filter(weather_vars =='ppt') %>%
  ggplot() +
  geom_boxplot(aes(x = clusters, y = values,fill = clusters)) +
  coord_flip() +
  facet_wrap(~weather_vars, ncol = 3,nrow = 3, scale ='free') +
  scale_fill_viridis_d(option = 'plasma') + 
  labs(fill = 'Clusters',title = 'Weather variables distribution across clusters')


ggsave(plot_summary_weather,filename = '../plots/plot_summary_weather.png',width = 10,height = 5)
plot_summary_weather
```

```{r}
plot_summary_soil <- df_grid_clusters %>%
  dplyr::select(c(clusters, 
                   bdod_0.5cm, sand_0.5cm, soc_0.5cm, 
                   silt_0.5cm, phh2o_0.5cm)) %>%
  pivot_longer(cols = -clusters,names_to = 'soil_vars',values_to = 'values') %>%
  mutate(clusters = as.factor(clusters)) %>%
  #filter(weather_vars =='ppt') %>%
  ggplot() +
  geom_boxplot(aes(x = clusters, y = values,fill = clusters)) +
  coord_flip() +
  facet_wrap(~soil_vars, ncol = 3,nrow = 3, scale ='free') +
  scale_fill_viridis_d(option = 'plasma') + 
  labs(fill = 'Clusters',title = 'Soil variables distribution across clusters')

ggsave(plot_summary_soil,filename = '../plots/plot_summary_soil.png',width = 10,height = 5)
plot_summary_soil
```



### 4.3 Crop data integration


```{r message=FALSE}
clusters_columns_df_crops <- raster::extract(round(cluster_raster$clusters), crop_data)

df_crop <- data.frame(crop_data, 
           raster::extract(weather_vars$aet, crop_data),
           raster::extract(weather_vars$tmax, crop_data),
           raster::extract(weather_vars$ppt, crop_data),
           raster::extract(weather_vars$tmin, crop_data),
           raster::extract(weather_vars$vpd, crop_data),
           
           raster::extract(soil_data_mask$`soc_0-5cm`, crop_data),
           raster::extract(soil_data_mask$`bdod_0-5cm`, crop_data),
           raster::extract(soil_data_mask$`sand_0-5cm`, crop_data),
           raster::extract(soil_data_mask$`silt_0-5cm`, crop_data),
           raster::extract(soil_data_mask$`phh2o_0-5cm`, crop_data),
           raster::extract(round(cluster_raster$clusters), crop_data)
           ) %>%
  dplyr::select(-c(ID.1,ID.2,ID.3,ID.4,ID.5,ID.6,ID.7,ID.8,ID.9)) %>%
  #na.exclude() %>%
  rename_with(.fn = ~ map_chr(., function(col_name) {
    str_replace(
      col_name,
      "^([a-zA-Z0-9]+)_\\d{4}\\.(\\d{2})\\.\\d{2}_\\d{8}$",
      "\\1_\\2"
    )
  })) 



cols_filtered <- names(df_crop) %>%
  keep(~ str_detect(., paste0("_(", paste(sprintf("%02d", c(1, 2, 3, 4, 11, 12)) , collapse = "|"), ")$")))

df_crop.X <- df_crop %>%
  st_as_sf(sf_column_name = "geometry") %>%
  st_set_crs(st_crs("epsg:4326")) %>%
  mutate(X = st_coordinates(.)[,1],
         Y = st_coordinates(.)[,2]) %>%
  st_transform(32720) %>%
  mutate(X_proj = st_coordinates(.)[,1],
         Y_proj = st_coordinates(.)[,2])  %>%
  dplyr::select(X_proj,Y_proj, Province, Trial_Site , GM, Cultivar, Empresa,Rendimiento,Env_index,
                all_of(cols_filtered),soc_0.5cm,bdod_0.5cm,sand_0.5cm,silt_0.5cm,phh2o_0.5cm,clusters) %>%
  mutate(clusters = as.factor(clusters)) %>%
  group_by(Cultivar) %>%
  mutate(Cultivar_id = as.factor(cur_group_id())) %>%
  ungroup()  %>% 
  mutate_if(is.character, as.factor) %>%
  rename(Yield = Rendimiento) %>%
  st_drop_geometry() %>%
  na.exclude()

```

```{r}
df_crop.X %>%
  ggplot() + 
  geom_smooth(aes(x = Env_index,y=Yield,color = Cultivar, group = Cultivar),
              alpha = 1/10,linewidth = 0.5,
              se = FALSE,method = 'lm',formula = 'y ~x', show.legend = FALSE) + 
  facet_wrap(~clusters)

```

```{r}
plot_summary_yield <- df_crop.X  %>%
  dplyr::select(c(clusters,GM, Yield)) %>%
  #pivot_longer(cols = -clusters,names_to = 'soil_vars',values_to = 'values') %>%
  mutate(clusters = as.factor(clusters)) %>%
  #filter(weather_vars =='ppt') %>%
  ggplot() +
  geom_boxplot(aes(x = clusters, y = Yield, fill = clusters)) +
  coord_flip() +
  #facet_wrap(~GM, ncol = 3,nrow = 3, scale ='free') +
  scale_fill_viridis_d(option = 'plasma') + 
  labs(fill = 'Clusters',title = 'Soybean seed yield  distribution across clusters')

ggsave(plot_summary_yield,filename = '../plots/plot_summary_yield.png',width = 10,height = 5)
plot_summary_yield
```




## 3. Machine Learning Workflow

### 3.1 Data Splitting

For robust model evaluation, we split the data into training (70%) and testing (30%) sets, stratified by the target variable.

```{r data_split}
set.seed(27)

df_split <- initial_split(df_crop.X, prop = 0.7,strata = 'clusters')
df_split
```

```{r extract_train_data}
df_train <- training(df_split)
df_train
```

```{r extract_test_data}

df_test <- testing(df_split)
df_test
```

### 3.2 Visualizing Data Distribution

Let's verify that our training and testing sets have similar distributions of the target variable.

```{r visualize_data_distribution}

plot_train <- df_train %>% mutate(Dataset = "Train")
plot_test <- df_test %>% mutate(Dataset = "Test")
combined_data <- bind_rows(plot_train, plot_test)

dist_plot_train_test <- ggplot(data = combined_data) +
  geom_rug(aes(x = Yield, fill = Dataset)) +
  geom_density(aes(x = Yield, fill = Dataset), color = NA, alpha = 0.3) +
  scale_fill_manual(values = c("red", "blue")) +
  labs(x = "Soybean Yield (kg/ha)", 
       y = "Density", 
       fill = "Dataset",
       title = "Distribution of Soybean Yield in Training and Testing Sets")

ggsave(dist_plot_train_test,filename = '../plots/dist_plot_train_test.png',width = 6,height = 3)
dist_plot_train_test
```
  
### 3.3 Feature Engineering

We define a recipe for data preprocessing, removing unnecessary columns.

```{r define_recipe}
df_recipe <-
  recipe(Yield ~ ., data = df_train) %>%

  step_rm(X_proj,Y_proj,Province,Trial_Site,GM,Cultivar,Empresa,Env_index) 

df_recipe
```

```{r prepare_recipe}
df_prep <- df_recipe %>% prep()
df_prep
```

## 4. Model Training

### 4.1 Model Specification

We define a Random Forest model with hyperparameters to be tuned.

```{r model_specification}
rf_spec <- 

  rand_forest(
    mtry = tune(),    
    trees = tune(),   
    min_n = tune()   
  ) %>% 
  set_engine("ranger", importance = "permutation") %>% 
  set_mode("regression")

rf_spec
```


### 4.2 Hyperparameter Tuning

We'll use 5-fold cross-validation to find the optimal hyperparameters.

```{r create_cv_folds}
set.seed(15)
resampling_foldcv <- vfold_cv(df_train, repeats = 3, v = 5)
```


```{r}

rf_param <- parameters(rf_spec) %>%
  recipes::update(
    mtry = mtry(c(1L, 100L)),
    trees = trees(c(100L, 4000L)),
    min_n = min_n(c(2L, 20L))
  )

set.seed(45)
registerDoParallel(cores = parallel::detectCores() - 1)

rf_grid_result <- tune_bayes(
  object = rf_spec,
  preprocessor = df_recipe,
  resamples = resampling_foldcv,
  param_info = rf_param,
  initial = 10,  
  iter = 30,     
  metrics = metric_set(rmse, rsq),
  control = control_bayes(
    no_improve = 10,
    verbose = TRUE,
    save_pred = TRUE
  )
)

stopImplicitCluster()
```



```{r}
rf_grid_result$.metrics[[1]]
```



### 4.3 Selecting Optimal Hyperparameters

We'll evaluate different strategies for selecting the best hyperparameters.

```{r select_by_rmse}

best_rmse <- rf_grid_result %>% 
  select_by_pct_loss("trees",
                     metric = "rmse",
                     limit = 2) %>% 
  mutate(source = "best_rmse")

best_rmse_2 <- rf_grid_result %>% 
  select_best(metric = "rmse") %>% 
  mutate(source = "best_rmse_2")

best_rmse_3 <- rf_grid_result %>% 
  select_by_one_std_err(metric = "rmse",
                        mtry) %>% 
  mutate(source = "best_rmse_3")
```

```{r select_by_r2}

best_r2 <- rf_grid_result %>% 
  select_by_pct_loss("trees",
                     metric = "rsq",
                     limit = 2) %>% 
  mutate(source = "best_r2")

best_r2_2 <- rf_grid_result %>% 
  select_best(metric = "rsq") %>% 
  mutate(source = "best_r2_2")

best_r2_3 <- rf_grid_result %>% 
  select_by_one_std_err(metric = "rsq",
                        mtry) %>%
  mutate(source = "best_r2_3")
```

```{r combine_hyperparameters}

hyperparameters_df <- best_rmse %>% 
  bind_rows(best_rmse_2, best_rmse_3, best_r2, best_r2_2, best_r2_3)
```

### 4.4 Comparing Hyperparameter Performance

We evaluate each set of hyperparameters on our validation data to select the best.

```{r hyperparameter_comparison_function}

compare_hyperparameters <- function(params_df, recipe, split) {

  results <- tibble(
    mtry = numeric(),
    trees = numeric(),
    min_n = numeric(),
    rmse = numeric(),
    source = character()
  )
  
  for(i in 1:nrow(params_df)) {

    current_params <- params_df[i, ]

    set.seed(10)
    current_spec <- rand_forest(
      mtry = current_params$mtry,
      trees = current_params$trees,
      min_n = current_params$min_n
    ) %>%
      set_engine("ranger", importance = "permutation") %>%
      set_mode('regression')

    current_fit <- last_fit(current_spec, recipe, split = split)
    current_metrics <- current_fit %>% collect_metrics()

    current_rmse <- current_metrics %>% 
      filter(.metric == "rmse") %>% 
      pull(.estimate)

    results <- results %>% add_row(
      mtry = current_params$mtry,
      trees = current_params$trees,
      min_n = current_params$min_n,
      rmse = current_rmse,
      source = current_params$source
    )
  }

  best_params <- results %>% 
    arrange(rmse) %>% 
    slice(1)
  
  return(list(
    all_results = results,
    best_combination = best_params
  ))
}

results <- compare_hyperparameters(
  params_df = hyperparameters_df,
  recipe = df_recipe,
  split = df_split
)

print(results)
```

### 4.5 Final Model Specification

Using the best hyperparameters, we specify our final model.

```{r final_model_specification}

final_spec <- rand_forest(
  trees = best_rmse$trees,
  mtry = best_rmse$mtry,
  min_n = best_rmse$min_n
) %>%
  set_engine("ranger", importance = "permutation") %>% 

  set_mode("regression")

final_spec
```

## 5. Model Validation

Now we fit the model on the training data and evaluate on the test data.

```{r final_model_fit}

set.seed(10)
final_fit <- last_fit(
  final_spec,
  df_recipe,
  split = df_split
)

final_fit %>%
  collect_predictions()
```

### 5.1 Test Set Performance

Let's examine how well our model performs on the unseen test data.

```{r test_metrics}

final_fit %>%
  collect_metrics()
```

### 5.2 Training Set Performance

For comparison, we also check performance on the training data to assess potential overfitting.

```{r train_metrics}

train_metrics <- final_spec %>%
  fit(Yield ~ .,
      data = bake(df_prep, df_train)) %>%
  augment(new_data = bake(df_prep, df_train))

train_metrics %>% 
  rmse(Yield, .pred) %>%
  bind_rows(
    train_metrics %>% 
    rsq(Yield, .pred)
  )
```

### 5.3 Prediction Visualization

Let's visualize how well our predictions match the observed values.

```{r predicted_vs_observed}

plot <- final_fit %>%
  collect_predictions() %>%
  metrica::scatter_plot(
    obs = Yield,
    pred = .pred,
    print_eq = TRUE,
    print_metrics = TRUE,
    metrics_list = c("R2", "RMSE", "KGE","NSE"),

    position_metrics = c(x=1000, y=6550),
    position_eq = c(x=1000, y=5000)
  )

ggsave(plot = plot,filename = '../plots/predict_vs_observed.png',width = 5,height = 5)


plot
```

### 5.4 Comprehensive Performance Metrics

Let's calculate a comprehensive set of performance metrics.

```{r comprehensive_metrics}

test_results <- final_fit %>% 
  collect_predictions()

metrics <- test_results %>% 
  summarise(
    RMSE = rmse_vec(truth = Yield, estimate = .pred),
    R2 = rsq_vec(Yield, .pred),
    Correlation = cor(.pred, Yield),
    SD = sd(.pred)  
  )

print(metrics)
```

## 6. Feature importance


```{r}

vip_df <- final_fit %>%
  extract_fit_parsnip() %>%
  vip::vi()  

var_importance <- vip_df %>%
  mutate(Variable = fct_reorder(Variable, Importance)) %>%  
  ggplot(aes(x = Variable, y = Importance)) +
  geom_col() +
  coord_flip() + 
  labs(title = 'Variable importance')

var_importance
ggsave(plot = var_importance,filename = '../plots/var_importance.png',width = 8,height = 5)

var_importance
```









